# netresponse - probabilistic tools for functional network analysis

For maintainer contact details, see the [README](../README.md) file


## Usage examples

Examples on running NetResponse algorithm and visualizing the
results. The algorithm combines network and functional information to
detect coherent subnetworks that reveal distinct activation modes
across conditions. Kindly cite [this
article](http://bioinformatics.oxfordjournals.org/content/26/21/2713.short).

```{r NetResponse1, warning=FALSE, message=FALSE}
library(netresponse)

# Generate simulated data
res <- generate.toydata()
D <- res$data
component.means <- res$means
component.sds   <- res$sds
sample2comp     <- res$sample2comp

# Fit NetResponse model
# Various network formats are supported, see help(detect.responses) for
# details. With large data sets, consider the 'speedup' option.
set.seed(4243)
res <- detect.responses(D, mixture.method = "vdp", pca.basis = TRUE)

# List subnets (each is a list of nodes)
subnet.id <- names(get.subnets(res))[[1]]
```

### PCA visualization

```{r NetResponse2, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "pca")
```

### Network visualization

```{r NetResponse3, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "network")
```

### Heatmap visualization

```{r NetResponse4, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "heatmap")
```

### Boxplot visualization

```{r NetResponse5, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "boxplot.data")
```

### Barplot visualization

```{r NetResponse6, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "response.barplot")
```

### Color scale

```{r NetResponse7, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot.scale(vis$breaks, vis$palette, two.sided = TRUE)
```




### Citing NetResponse

Please cite [Lahti et al. (2010)](http://bioinformatics.oxfordjournals.org/content/26/21/2713) with the package. When using the ICMg algorithms, additionally cite [Parkkinen et al. (2010)](http://www.biomedcentral.com/1752-0509/4/4).


### Version information

This document was written using:

```{r version, warning=FALSE, message=FALSE}
sessionInfo()
```
