# netresponse - probabilistic tools for functional network analysis

For maintainer contact details, see the [README](../README.md) file

## Background 

Condition-specific network activation is characteristic for cellular
systems and other real-world interaction networks. If measurements of
network states are available across a versatile set of conditions or
time points, it becomes possible to construct a global view of network
activation patterns. Different parts of the network respond to
different conditions, and in different ways. Systematic, data-driven
identification of these responses will help to obtain a holistic view
of network activity \cite{Lahti10bioinf, Lahti10thesis}. This package
provides robust probabilistic algorithms for functional network
analysis \cite{Lahti10bioinf, Parkkinen10bmcsysbio}. 

The methods are based on nonparametric probabilistic modeling and
variational learning, and provide general exploratory tools to
investigate the structure (ICMg; \cite{Parkkinen10bmcsysbio}) and
context-specific behavior (NetResponse; \cite{Lahti10bioinf}) of
interaction networks.  ICMg is used to identify community structure in
interaction networks; NetResponse detects and characterizes
subnetworks that exhibit context-specific activation patterns across
versatile collections of functional measurements, such as gene
expression data. The implementations are partially based on the
agglomerative independent variable group analysis \citep{Honkela08}
and variational Dirichlet process Gaussian mixture models
\cite{Kurihara07nips}. The tools are particularly useful for global
exploratory analysis of genome-wide interaction networks and versatile
collections of gene expression data.


## Usage examples

Examples on running NetResponse algorithm and visualizing the
results. The algorithm combines network and functional information to
detect coherent subnetworks that reveal distinct activation modes
across conditions. Kindly cite [this
article](http://bioinformatics.oxfordjournals.org/content/26/21/2713.short).

```{r NetResponse1, warning=FALSE, message=FALSE}
library(netresponse)

# Generate simulated data
res <- generate.toydata()
D <- res$data
component.means <- res$means
component.sds   <- res$sds
sample2comp     <- res$sample2comp

# Fit NetResponse model
# Various network formats are supported, see help(detect.responses) for
# details. With large data sets, consider the 'speedup' option.
set.seed(4243)
res <- detect.responses(D, mixture.method = "vdp", pca.basis = TRUE)

# List subnets (each is a list of nodes)
subnet.id <- names(get.subnets(res))[[1]]
```

### PCA visualization

```{r NetResponse2, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "pca")
```

### Network visualization

```{r NetResponse3, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "network")
```

### Heatmap visualization

```{r NetResponse4, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "heatmap")
```

### Boxplot visualization

```{r NetResponse5, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "boxplot.data")
```

### Barplot visualization

```{r NetResponse6, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "response.barplot")
```

### Color scale

```{r NetResponse7, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot.scale(vis$breaks, vis$palette, two.sided = TRUE)
```


### Cluster assignments

The sample-response assignments from the mixture model are soft
ie. defined as continuous probabilities. Retrieve the hard clustering
ie. list of samples for each response, response for each sample, based
the highest probability:

```{r NetResponse8, warning=FALSE, message=FALSE}
subnet.id <- 'Subnet-2'
sample.probs <- response2sample(model, subnet.id)
response.probs <- sample2response(model, subnet.id)
```

Retrieve model parameters for a given subnetwork (Gaussian mixture
means, covariance diagonal, and component weights):

```{r NetResponse9, warning=FALSE, message=FALSE}
get.model.parameters(model, subnet.id) 
```






### Citing NetResponse

Please cite [Lahti et al. (2010)](http://bioinformatics.oxfordjournals.org/content/26/21/2713) with the package. When using the ICMg algorithms, additionally cite [Parkkinen et al. (2010)](http://www.biomedcentral.com/1752-0509/4/4).


### Version information

This document was written using:

```{r version, warning=FALSE, message=FALSE}
sessionInfo()
```
