# NetResponse examples

Examples on running NetResponse algorithm and visualizing the
results. The algorithm combines network and functional information to
detect coherent subnetworks that reveal distinct activation modes
across conditions. Kindly cite [this
article](http://bioinformatics.oxfordjournals.org/content/26/21/2713.short).

Acknowledgements: This program is based on the AIVGA Agglomerative
Independent Variable Group Analysis package (v. 1.0) Copyright (C)
2001-2007 Esa Alhoniemi, Antti Honkela, Krista Lagus, Jeremias
Seppa, Harri Valpola, and Paul Wagner.

```{r NetResponse1, warning=FALSE, message=FALSE}
library(netresponse)

# Generate simulated data
res <- generate.toydata()
D <- res$data
component.means <- res$means
component.sds   <- res$sds
sample2comp     <- res$sample2comp

# Fit NetResponse model
set.seed(4243)
res <- detect.responses(D, mixture.method = "vdp", pca.basis = TRUE)

# List subnets (each is a list of nodes)
subnet.id <- names(get.subnets(res))[[1]]
```

### PCA visualization

```{r NetResponse2, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "pca")
```

### Network visualization

```{r NetResponse3, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "network")
```

### Heatmap visualization

```{r NetResponse4, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "heatmap")
```

### Boxplot visualization

```{r NetResponse5, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "boxplot.data")
```

### Barplot visualization

```{r NetResponse6, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
vis <- plot.responses(res, subnet.id, plot.mode = "response.barplot")
```

### Color scale

```{r NetResponse7, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot.scale(vis$breaks, vis$palette, two.sided = TRUE)
```
